Reanalysis Step 2: Multiple Imputation (Clean Version)

### 2025-08-29 reanalysis-2-multiple_imputation_clean.qmd
refactor

```{r prepare-data}
# Select variables and clean up
dr_cols <- c("PDR", "NPDR", "No_DR")
for (col in setdiff(dr_cols, names(df_original))) {
  df_original[[col]] <- NA_integer_
}
if (!"person_dr" %in% names(df_original)) {
  df_original[["person_dr"]] <- NA_integer_
}
df_mice <- df_original |>
  select(all_of(all_vars), any_of(dr_cols)) |>
  mutate(
    # Create person_dr from DR columns if needed
    person_dr = case_when(
      !is.na(PDR) & PDR == 1 ~ 2L,
      !is.na(NPDR) & NPDR == 1 ~ 1L,
      !is.na(No_DR) & No_DR == 1 ~ 0L,
      TRUE ~ person_dr
    ),
    # Convert to ordered factor
    person_dr = factor(person_dr, levels = c(0L, 1L, 2L), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Recode comorbidities to 0/1
    Glaucoma_bef_hitplus = ifelse(Glaucoma_bef_hitplus == "Present", 1, 0),
    Otherretina_bef_hitplus = ifelse(Otherretina_bef_hitplus == "Present", 1, 0),
    # Clean factor levels (remove spaces for formula compatibility)
    across(where(is.character), ~ factor(gsub("[^[:alnum:]_]", "_", .x)))
  )

@@

@@

>>>>

```{r prepare-data}
# Select variables and clean up
dr_cols <- c("PDR", "NPDR", "No_DR")
for (col in setdiff(dr_cols, names(df_original))) {
  df_original[[col]] <- NA_integer_
}
if (!"person_dr" %in% names(df_original)) {
  df_original[["person_dr"]] <- NA_integer_
}
df_mice <- df_original |>
  select(any_of(all_vars), any_of(dr_cols)) |>
  mutate(
    # Create person_dr from DR columns if needed
    person_dr = case_when(
      !is.na(PDR) & PDR == 1 ~ 2L,
      !is.na(NPDR) & NPDR == 1 ~ 1L,
      !is.na(No_DR) & No_DR == 1 ~ 0L,
      TRUE ~ person_dr
    ),
    # Convert to ordered factor
    person_dr = factor(person_dr, levels = c(0L, 1L, 2L), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Recode comorbidities to 0/1
    Glaucoma_bef_hitplus = ifelse(Glaucoma_bef_hitplus == "Present", 1, 0),
    Otherretina_bef_hitplus = ifelse(Otherretina_bef_hitplus == "Present", 1, 0),
    # Clean factor levels (remove spaces for formula compatibility)
    across(where(is.character), ~ factor(gsub("[^[:alnum:]_]", "_", .x)))
  ) |>
  dplyr::group_by(e_mrn_deidentified) |>
  dplyr::mutate(
    n_eyes = dplyr::n(),
    eye_index = dplyr::row_number(),
    fellow_baseline_VA = dplyr::if_else(n_eyes == 2, baseline_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_intermediate_VA = dplyr::if_else(n_eyes == 2, intermediate_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_outcome_VA = dplyr::if_else(n_eyes == 2, outcome_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_dr = dplyr::if_else(n_eyes == 2,
                               person_dr[3 - eye_index],
                               factor(NA, levels = levels(person_dr), ordered = TRUE))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(-eye_index)

@@

# Fellow-eye variables: predictors only (do not impute)
vars_fellow <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr", "n_eyes")
for (var in vars_fellow) {
  if (var %in% names(methods)) {
    methods[var] <- ""
  }
}

@@

# Fellow-eye predictors
fellow_vars <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr")
for (v in fellow_vars) {
  if (v %in% rownames(pred_matrix)) {
    pred_matrix[v, ] <- 0
  }
}
if ("fellow_baseline_VA" %in% colnames(pred_matrix)) {
  pred_matrix["baseline_VA_logMAR", "fellow_baseline_VA"] <- 1
}
if ("fellow_intermediate_VA" %in% colnames(pred_matrix)) {
  pred_matrix["intermediate_VA_logMAR", "fellow_intermediate_VA"] <- 1
}
if ("fellow_outcome_VA" %in% colnames(pred_matrix)) {
  pred_matrix["outcome_VA_logMAR", "fellow_outcome_VA"] <- 1
}
if ("fellow_dr" %in% colnames(pred_matrix) && "person_dr" %in% rownames(pred_matrix)) {
  pred_matrix["person_dr", "fellow_dr"] <- 1
}
